---
title: "R Notebook"
output: html_notebook
---

```{r}
library(rpart)
library(rpart.plot)
library(tidyverse)

library(tidyverse)

```



```{r}
titanic_set <- read_csv("data/titanic_decision_tree_data.csv")
```

```{r}
shuffle_index <- sample(1:nrow(titanic_set))

# shuffle the data so class order isn't in order - need this for training/testing split later on 
titanic_set <- titanic_set[shuffle_index, ]
```

```{r}
head(titanic_set)
```

1.1

```{r}
titanic_clean <- titanic_set %>%
  filter(survived %in% c(0,1)) %>%
# Convert to factor level
    mutate(sex = as.factor(sex), 
           age_status = as.factor(if_else(age <= 16, "child", "adult")),
         class = factor(pclass, levels = c(3,2,1), labels = c("Lower", "Middle", "Upper")), 
           survived_flag = factor(survived, levels = c(0,1), labels = c("No", "Yes")), 
           port_embarkation = as.factor(embarked)) %>%
  select(sex, age_status, class, port_embarkation, sib_sp, parch, survived_flag) %>%
  na.omit()
```

1.2 Question 2

Have a look at your data and create some plots to ensure you know what you’re working with before you begin. Write a summary of what you have found in your plots. Which variables do you think might be useful to predict whether or not people are going to die? Knowing this before you start is the best way to have a sanity check that your model is doing a good job.

```{r}
library(GGally)
```

```{r}
ggpairs(titanic_clean)
```

1.3 Question 3

Now you can start to build your model. Create your testing and training set using an appropriate split. Check you have balanced sets. Write down why you chose the split you did and produce output tables to show whether or not it is balanced.

```{r}
n_data <- nrow(titanic_clean)

# created a text index
test_index <- sample(1:n_data, size = n_data*0.2)

#created test set

titanic_test <- slice(titanic_clean, test_index)

#created training data

titanic_train <- slice(titanic_clean, -test_index)
```

```{r}
library(janitor)
```

```{r}
titanic_test %>%
  tabyl(survived_flag)
```

```{r}
titanic_train %>%
  tabyl(survived_flag)
```

These look pretty balanced datasets to work with, i chose the 20 percent as i thought it was a good starting point, given the size of the dataset

1.4 Question 4

Create your decision tree to try and predict survival probability using an appropriate method, and create a decision tree plot.

```{r}
titanic_fit <- rpart(survived_flag ~ ., 
                     data = titanic_train,
                     method = "class")

titanic_fit
```

```{r}
rpart.plot(titanic_fit, yesno = 2)
```

This has decided that sex = male as the most important element, the important variable are
sex = male, class = lower, middle, age_status = adult, slb_sp >= 3, parch < 1, slb_sp < 1, class = lower, slb_sp >= 1.

the highest chance of survival is the if the sex = female. the lowest chance of not surviving is class = lower, middle

1.5 Question 6

Test and add your predicitons to your data. Create a confusion matrix. Write down in detial what this tells you for this specific dataset.

```{r}
library(modelr)
```

```{r}
titanic_test_pred <- titanic_test %>% 
    add_predictions(titanic_fit, type = 'class')
```

```{r}
# look at the variables 

titanic_test_pred %>% 
  select(sex, class, age_status, sib_sp, parch, pred)
```

```{r}
rpart.predict(titanic_fit, newdata=titanic_test[1:3,], rules=TRUE)
```

```{r}
library(yardstick)
```

```{r}
conf_mat <- titanic_test_pred %>%
              conf_mat(truth = survived_flag, estimate = pred)

conf_mat

```


This shows the the more accurate the tree the higher the diagonal will be. This should be pretty accurate

```{r}
accuracy <- titanic_test_pred %>%
  accuracy(truth = survived_flag, estimate = pred)
```

```{r}
library(caret)

```

```{r}
confusionMatrix(titanic_test_pred$pred, titanic_test_pred$survived_flag) 
```


CODECLAN_SOLUTION

```{r}
titanic_clean_1 <- titanic_set %>%
  filter(survived %in% c(0,1)) %>%
# Convert to factor level
    mutate(sex = as.factor(sex), 
           age_status = as.factor(if_else(age <= 16, "child", "adult")),
         class = factor(pclass, levels = c(3,2,1), labels = c("Lower", "Middle", "Upper")), 
           survived_flag = factor(survived, levels = c(0,1), labels = c("dies", "survived")), 
           port_embarkation = as.factor(embarked)) %>%
  select(sex, age_status, class, port_embarkation, sib_sp, parch, survived_flag) %>%
  na.omit()

```


```{r}
titanic_clean_1 %>%
  select(survived_flag, sex) %>%
  group_by(survived_flag,sex) %>%
  summarise(total_count = n()) %>%
ggplot() + 
  aes(x = survived_flag, y = total_count, fill = sex) + 
  geom_bar(stat = "identity", position = position_dodge()) + 
  scale_fill_brewer(palette="Dark2") + 
  ylab("total count") + xlab("survived")

```


```{r}
titanic_clean_1 %>%
  select(survived_flag, class) %>%
  group_by(survived_flag,class) %>%
  summarise(total_count = n()) %>%
ggplot() + 
  aes(x = survived_flag, y = total_count, fill = class) + 
  geom_bar(stat = "identity", position = position_dodge()) + 
  scale_fill_brewer(palette="Dark2") + 
  ylab("total count") + xlab("survived")
```

```{r}
titanic_clean_1 %>%
  select(survived_flag, age_status) %>%
  group_by(survived_flag,age_status) %>%
  summarise(total_count = n()) %>%
ggplot() + 
  aes(x = survived_flag, y = total_count, fill = age_status) + 
  geom_bar(stat = "identity", position = position_dodge()) + 
  scale_fill_brewer(palette="Dark2") + 
  ylab("total count") + xlab("survived")
```

```{r}
titanic_clean_1 %>%
  select(survived_flag, port_embarkation) %>%
  group_by(survived_flag,port_embarkation) %>%
  summarise(total_count = n()) %>%
ggplot() + 
  aes(x = survived_flag, y = total_count, fill = port_embarkation) + 
  geom_bar(stat = "identity", position = position_dodge()) + 
  scale_fill_brewer(palette="Dark2") + 
  ylab("total count") + xlab("survived")
```

```{r}
# set the random seed number: ONLY INSTRUCTORS - students SHOULD NOT do this in their homework 
set.seed(3)

n_data_1 <- nrow(titanic_clean_1)

# create a test sample index
test_index_1 <- sample(1:n_data, size = n_data*0.2)

# create test set
titanic_test_1  <- slice(titanic_clean_1, test_index)

# create training set
titanic_train_1 <- slice(titanic_clean_1, -test_index)

# check balanced sets
titanic_test_1 %>%
 janitor::tabyl(survived_flag)
```

```{r}
titanic_train_1 %>%
  janitor::tabyl(survived_flag)

```


```{r}
titanic_fit_1 <- rpart(survived_flag ~ ., 
                     data = titanic_train_1, 
                     method = 'class')

rpart.plot(titanic_fit_1, yesno = 2)
```

```{r}
library(modelr)
library(yardstick)
```

```{r}
# add the predictions
titanic_test_pred_1 <- titanic_test_1 %>%
                 add_predictions(titanic_fit_1, type = 'class')


conf_mat <- titanic_test_pred_1 %>%
              conf_mat(truth = survived_flag, estimate = pred)

conf_mat
```

EXTENSION

```{r}
control <- rpart.control(minsplit = 4,
    minbucket = round(5 / 3),
    maxdepth = 3,
    cp = 0)

tune_fit <- rpart(survived_flag ~., data = titanic_train, method = 'class', control = control)

rpart.plot(tune_fit, yesno = 2)
```

