---
title: "R Notebook"
output: html_notebook
---



```{r}
library(RPostgreSQL)

```


```{r}

home_location <- Sys.getenv("HOME")
credentials_location <- paste0(home_location,"/credentials.R")
source(credentials_location)

```


Create a connection to the local wrestling database from R.
Pass the argument max.con = 2 to PostgreSQL() to make sure you don’t have too many simultaneous connections. We’ll be setting up a local connection and perhaps a remote connection, hence ‘2’.
Remember to add a separate ‘disconnect’ block of code. We’ll write queries between the ‘connect’ and ‘disconnect’ blocks.


```{r}
## This will seed the database TO r

db_connect_2 <- dbConnect(drv = PostgreSQL(max.con = 2), user = postgresql_username, password = postgresql_password, dbname = 'wrestling', host = 'localhost')

```

Using the R connection you just created, write and execute a SQL query to return all the wrestlers in the wrestlers table.


```{r}
dbListTables(conn = db_connect_2)

```

```{r}
## this provides the fields that are in the table so to know which colunn to call in the next code
dbListFields(conn = db_connect_2, name = "wrestlers")
```


```{r}

## this will pull all the details in the table

wrestlers <- dbGetQuery(conn = db_connect_2, statement = "SELECT * FROM wrestlers")

wrestlers

```


Explore the columns stored in the table. State which data type you think each column stores.


id = INT, born = VARCHAR, name- VARCHAR, gender- VARCHAR, win_percentage- REAL, total_matches- INT, 


```{r}
dbDisconnect(conn = db_connect_2)


```


Now we’ll try to connect to a remote PostgreSQL database holding the same data. As earlier, this connection may or may not work, depending on how many other connections to the server are currently active. But try a few times, and don’t worry if it proves difficult. Being able to write the connection code is what we’re looking for here!

Here are the details of the remote PostgreSQL database we wish to connect to:


  

```{r}
##this will course in the credential location

source(credentials_location)

```


```{r}

## this is the ocde to pull the database from the rmote location

db_connect <- dbConnect(drv = PostgreSQL(max.con = 2), user = remote_wrestling_postgresql_username, password = remote_wrestling_postgresql_password, dbname='mvhbdgdk', host='34.245.37.126')

```


```{r}
db_connect

```

Using this R connection, write and execute a query to return all the wrestlers from the wrestlers table

```{r}
## This code pulls the database information from the relevant remote connection

result <- dbGetQuery(conn = db_connect, statement = "SELECT * FROM wrestlers")
result

```


```{r}
## this will disconnest the current remote connection

dbDisconnect(conn = db_connect)

```




