---
title: "Mini-Lab - Data Quality Answers"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    css: ../../../styles.css
  pdf_document: default
---

# Introduction

Now is a chance to practice defining, documenting and coding our own data quality rules.

Have a look at the `contractsregister.csv` file which comes from [Aberdeen Council's open data portal](https://www.aberdeenshire.gov.uk/online/open-data/).

# MVP - Defining and coding data quality rules
**Task - 1 hour**

## Business Statement

We are going to start by writing 3 business statements. Ensure they explain the required rule and the context as to why this is important for the business.

1. Write a statement ensuring the DateOfAward is populated

2. Write a statement ensuring that DateOfAward is in the past and is formatted correctly

3. Write a statement ensuring that the ContractValue is a positive numeric value

<details>
<summary>**Answers**</summary>

1. The date of award cannot be empty as this is a mandatory system field

2. The date of award should be a valid date formatted dd/mm/yyyy. The value should be either today's date or in the past. This field is used to identify when contracts require review.

3. The contract value must be populated with a positive numeric value. Negative or zero contract values are not valid.

</details>

## Rule Specification

Now for each business statement write rule(s) explaining the test that is being carried out. Ensure you can map each rule to the DQ dimension it is checking.

<details>
<summary>**Hint**</summary>

You may need more than one rule per business statement

</details>

<details>
<summary>**Answers**</summary>

1a. DateOfAward not empty or equal to "(blank)" - **completeness**

2a. DateOfAward should be formatted as 'dd/mm/yyyy' and should be a valid date - **validity**

2b. DateOfAward should be today's date or in the past - **validity**

3a. The ContractValue is not empty - **completeness**

3b. The ContactValue is greater than zero - **validity**

</details>

## The coded rules

For each data quality rule defined above, write `R-code` that can be used to test that rule.

You may want to use this function, which checks if a date is invalid:

```{r}
is_invalid_date <- function(date){
 
  parsed_date <- try(as.Date(date, format= "%d/%m/%Y"))
  
  return(inherits(parsed_date, "try-error") | is.na(parsed_date))
}
```

And this function, which checks if a date is in the past:

```{r}
is_in_past <- function(date){
  
  return(lubridate::dmy(date) <= lubridate::today())
  
}
```

Don't worry if you don't understand the code! We'll learn about dates and times in week 3.

<details>
<summary>**Answers**</summary>

```{r}
library(readr)
library(lubridate)

contracts <- read_csv("data/contractsregister.csv")

rule1a <- contracts$DateOfAward == "(blank)" |
  contracts$DateOfAward == "" |
  is.na(contracts$DateOfAward)

rule2a <- is_invalid_date(contracts$DateOfAward)

rule2b <- rule2a | !is_in_past(contracts$DateOfAward)

rule3a <- is.na(contracts$ContractValue)

rule3b <- contracts$ContractValue <= 0  
```


</details>


## Summary statistics 

Create summary statistics, showing the volumes that fail each of the rules you have defined.

<details>
<summary>**Answers**</summary>

```{r}
sum(rule1a)
sum(rule2a)
sum(rule2b)
sum(rule3a)
sum(rule3b)
```

</details>







