---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(outliers)
```


1. Load the code_clan_tweets.csv data. Find the number of rows, columns, and list all the variable names.

```{r}
cc_tweets <- read_csv("data/code_clan_tweets.csv")

```


```{r}
dim(cc_tweets)

names(cc_tweets)

head(cc_tweets)

view(cc_tweets)
```

2. First, we’re going to summarise the number of likes (favorite_count) CodeClan tweets have. Create a boxplot to check for outliers.

```{r}
ggplot(cc_tweets, aes(x = "cc_tweets", y = "favourite_count")) +
  geom_boxplot() + ylab('favourite_count \n') +  xlab("\n cc tweets")

```




3. Find the total number of favourited tweets (stored in favorite_count) that CodeClan tweets have got. Don’t add a tweet’s favorite_count to the total when the tweet was quoted (see the is_quote variable).




Summarise the mean number of retweets (retweet_count), split by the type of platform that was used to tweet (source). Again, omit quoted tweets from the mean.

```{r}
summary(cc_tweets$retweet_count)

```

```{r}

cc_tweets %>%
  group_by(retweet_count) %>%
  summarise(n = n(), average = mean(retweet_count))
            

```

```{r}

retweet_data_mean <- cc_tweets %>%
  summarise(count = sum(retweet_count))
retweet_data_mean

```



4. Count the total number of likes (i.e. total of favorite_count), split by media type, and arrange them from most likes to least. Assume that any tweets without a listed media type are of type “text”.



5. Find the mean number of characters that a CodeClan tweet contains.



6. The code_clan_info.csv data file contains status_url along with other info. Load this in, and join it to the code_clan_tweets tibble, so that you have a status_url for each tweet. Decide which variable to join the tibbles on.

```{r}
codeclan_info <- read_csv("data/code_clan_info.csv")

```


```{r}
inner_join(cc_tweets, codeclan_info, "tweet_id")

```


7. From your new joined data, create a new tibble codeclan_hashtags containing only the tweet_id and hashtags in lowercase for analysis. Keep only those tweets with hashtags.