---
title: "R Notebook"
output: html_notebook
---

```{r}
library(dplyr)
library(tidyr)
library(tidytext)
library(janeaustenr)
```

Find the most common words in both Pride & Prejudice and Sense & Sensibility.

```{r}
pride_p <- tibble(
  sentence = 1:length(prideprejudice), 
  text = prideprejudice
) %>% 
  unnest_tokens(word, text) 

pride_p %>%
  count(word, sort = TRUE)
```


```{r}
sense_s <- tibble(
  sentence = 1:length(sensesensibility),
  text = sensesensibility
) %>% 
  unnest_tokens(word, text)

sense_s %>%
  count(word, sort = TRUE)
```

ind the most common words in both Pride & Prejudice and Sense & Sensibility, not including stop words

```{r}
pride_p %>%
  anti_join(stop_words) %>%
  count(word, sort = TRUE)

```

```{r}
sense_s %>%
  anti_join(stop_words) %>%
  count(word, sort = TRUE)
```

Find the most common sentiment words in both Pride & Prejudice and Sense & Sensibility.

```{r}
library(textdata)
```

```{r}
  pride_p_sentiment <- pride_p %>%
  inner_join(get_sentiments("afinn")) %>%
  arrange(desc(value))
```

```{r}
sense_s_sentiment <- sense_s %>%
inner_join(get_sentiments("afinn")) %>%
arrange(desc(value))

```

CODECLAN SOLUTION

```{r}
austen_words <- 
austen_books() %>%
  filter(book == "Pride & Prejudice" | book == "Sense & Sensibility") %>%
  unnest_tokens(word, text) %>%
  count(word, book, sort = TRUE)
```

```{r}
austen_words %>%
  filter(book == "Pride & Prejudice")
```

```{r}
austen_words %>%
  filter(book == "Sense & Sensibility")
```

```{r}
# 2
austen_words <- 
austen_words %>%
  anti_join(stop_words)
```

```{r}
austen_words %>%
  filter(book == "Pride & Prejudice")
```

```{r}
austen_words %>%
  filter(book == "Sense & Sensibility")
```

```{r}
# 3
austen_sentiment_words <-
  austen_words %>%
  inner_join(get_sentiments("bing"))
```

```{r}
austen_sentiment_words %>%
  filter(book == "Pride & Prejudice")
```

```{r}

austen_sentiment_words %>%
  filter(book == "Sense & Sensibility")

```

2 Extension

Taking your results above. Can you create a plot which visualises the differences between the books?

```{r}
library(ggplot2)

austen_words <-
austen_words %>%
mutate(
  sentiment = "Neutral"
)
  
plot_df <- 
rbind(
  slice(austen_words, 1:10),
  slice(austen_sentiment_words, 1:10)
)

ggplot(plot_df) +
  aes(x = word, y = n, fill = book) +
  geom_col(position = "dodge") +
  coord_flip() +
  facet_wrap(sentiment ~ ., scales = "free")
```

