---
title: "R Notebook"
output: html_notebook
---

```{r}
library(readr)
```


1. Read in the state_income_data.csv file usine the approprial function


```{r}  

## This reads in the csv file using

income_data <- read_csv("data/state_income_data.csv")

```



```{r}
## imported data

income_data
```


```{r}

class(income_data)

View(income_data)
```


2. How do you specify your own column names when using read_csv?

```{r}

## thisa is an example of the code that would be used to speicy own column names
 
income_data <- read.csv("data/state_income_data.csv", skip = 1, 
                        col_names = c( "Index",
                                       "state", 
                                       "y2002"
                                      ))           
                                

```

3. read in cars_csv file and use the col_types() function to ensure you load in x1 as char, mpg, cyl, disp, hp as double and vs as logical.


```{r}
## Loaded in thge csv file

cars_data <- read_csv("data/cars.csv")

```

```{r}

cars_data

```


```{r}
class(cars_data)

```




```{r}
## code input to change the csv file, so the vs column became logical

cars_data <- read_csv("data/cars.csv", col_types = cols(vs = col_logical()))

```


```{r}

## vs column now showing logicAL
cars_data

```


4. read the Scotland.txt file using the appopriate function


```{r}

## This is the function to read text file

scotland_text <- read_file("data/scotland.txt")

class(scotland_text)

scotland_text
```

5. How do you read text into one element character vector?

```{r}
## this read the text as one element

scotland_two <- read_file("data/scotland.txt")

scotland_two
```


6. Read in the steam_200k_sample.csv file. fix any problems you may encounter



```{r}

## when loading the code this error is received due to errors in the first 2 rows of the data

steam_200k_data <- read.csv("data/steam-200k-sample.csv")

```

```{r}

## This will skip the first two rows when it reads the data in

steam_200k_data <- read.csv("data/steam-200k-sample.csv", skip = 2)

steam_200k_data
```

7. read in the historic_eu_payments.tsv file. 

```{r}
## This is how you read tsv file in

historic <- read_tsv("data/historic_eu_payments.tsv")
```


8. How do you read lines of text into a multi element character vector

```{r}

## this is to reead line of text into multi elemnt character vector
  ## this is done usong the delim code e.g

beer <- read_delim("data/beer.txt", delim = ";")

```



9. Read in the JSON file from the following url: http://makemeanalyst.com/wp-content/uploads/2017/06/sample_json.txt


```{r}

library(jsonlite)
library(readr)

```



```{r}
## This is how to read in a JSON file

raw_makeme <- read_json("http://makemeanalyst.com/wp-content/uploads/2017/06/sample_json.txt")

```

```{r}

## This is the Json file
raw_makeme

```

```{r}

class(raw_makeme)
```
