---
title: "R Notebook"
output: html_notebook
---

```{r}
library(janitor)
library(fastDummies)
library(broom)
library(rpart)
library(rpart.plot)
library(tidyverse)
library(factoextra)
```

```{r}
customer_data <- read_csv("data/mall_customers.csv")
```

```{r}
customer_data <- customer_data %>%
  clean_names()
```

```{r}
summary(customer_data)
```

# find the data to be quite straight forward and should be reaonable to use for clustering purposes as there is only a few variables, there is no NA's either

We are interested in creating a marketing campaign to target customers based on their spending score and annual income. Perform a k-means clustering to find if there are meaningful clusters in the data to target the customers.

```{r}
ggplot(customer_data, aes(x = spending_score_1_100, y = annual_income_k)) +
  geom_point()
```


Perform k-means clustering and chose a value of k.

```{r}
customer <- customer_data %>%
  select(spending_score_1_100, annual_income_k)

customer
```

```{r}
summary(customer)
```

```{r}
customer_scale <- customer %>%
  mutate_all(scale)

customer_scale
```

```{r}
clustered_customers <- kmeans(customer_scale, centers = 6, nstart = 25)
clustered_customers
```

```{r}
library(broom)
```

```{r}
tidy(clustered_customers)

```

```{r}
glance(clustered_customers)
```


```{r}
augment(clustered_customers, customer)
```

```{r}
library(animation)

customer_scale %>% 
  kmeans.ani(centers = 6)
```


Visualise the clustering for your chosen value of k.

```{r}
# Set min & max number of clusters want to look at 
max_k <- 20 

k_clusters <- tibble(k = 1:max_k) %>%
  mutate(
    kclust = map(k, ~ kmeans(customer_scale, .x, nstart = 25)), 
    tidied = map(kclust, tidy),
    glanced = map(kclust, glance),
    augmented = map(kclust, augment, customer)
  )

k_clusters

```

```{r}
clusterings <- k_clusters %>%
  unnest(glanced)

clusterings
```

```{r}
ggplot(clusterings, aes(x=k, y=tot.withinss)) +
  geom_point() +
    geom_line() +
    scale_x_continuous(breaks = seq(1, 20, by = 1))
```

```{r}
fviz_nbclust(customer_scale, kmeans, method = "wss", nstart = 25)
```

```{r}
fviz_nbclust(customer_scale, kmeans, method = "silhouette", nstart = 25)
```

```{r}
fviz_nbclust(customer_scale, kmeans, method = "gap_stat", nstart = 25, k.max = 10)
```

```{r}
 clusterings %>% 
  unnest(cols = c(augmented)) %>%
  filter(k <= 5) %>%
 ggplot(aes(x = spending_score_1_100, y = annual_income_k)) +
  geom_point(aes(color = .cluster)) + 
  facet_wrap(~ k)
```

```{r}
 clusterings %>% 
  unnest(cols = c(augmented)) %>%
  filter(k == 5) %>%
 ggplot(aes(x = spending_score_1_100, y = annual_income_k, colour = .cluster)) +
  geom_point(aes(color = .cluster)) 
```

```{r}
 clusterings %>% 
  unnest(augmented) %>%
  filter(k == 5) %>%
  group_by(.cluster) %>%
  summarise(mean(spending_score_1_100), mean(annual_income_k))
```


Do you think the clustering seems a good fit for this data?

# Yes this clustering seems good for this data, at a k means = 5, a cluster of 5 seems to fit well.

Comment on the attributes on one or two of the clusters (maybe even give them a label if you like - like in section 4.1 of the ‘Segmentation & clustering intro’ lesson).

# The clusters shouw quite a close spread of the clustered data the purple cluster is very close clsutered.

CODECLAN SOLUTION

```{r}
library(tidyverse)
library(janitor)
```

```{r}
customer_data %>%
  group_by(gender) %>%
  summarise(count = n()) %>%
ggplot(aes(x = gender, y = count)) +
  geom_col()
```

```{r}
ggplot(customer, aes(x = annual_income_k, y = spending_score_1_100)) +
  geom_point()
```

```{r}
ggplot(customer, aes(x = annual_income_k)) +
  geom_histogram()
```

```{r}
ggplot(customer, aes(x = spending_score_1_100)) +
  geom_histogram()
```

```{r}
ggplot(customer_data, aes(x = age)) +
  geom_histogram(bins = 50)
```

```{r}
customers_scale_1 <- customer %>% 
  select(annual_income_k, spending_score_1_100) %>%
    mutate_all(scale)

summary(customers_scale_1)
```

```{r}
ggplot(customers_scale_1, aes(x = annual_income_k, y = spending_score_1_100)) +
  geom_point()
```

```{r}
customer_clusters_k5 <- k_clusters %>% 
  unnest(cols = c(augmented)) %>%
  filter(k == 5) 
```

