---
title: "R Notebook"
output: html_notebook
---

```{r}

library(RPostgreSQL)

```

```{r}

home_location <- Sys.getenv("HOME")
credentials_location <- paste0(home_location,"/credentials.R")
source(credentials_location)

db_connect_3 <- dbConnect(drv = PostgreSQL(), user = postgresql_username, password = postgresql_password, dbname = "acme_employees", host = "localhost")
rm(postgresql_username, postgresql_password)


```

Create a new data.frame containing the details of all employees held in the database


```{sql connection=db_connect_3}

SELECT * 
FROM employees AS e
INNER JOIN Teams AS T
ON e.id = t.id


```

Retrieve the records for all employees again, but this time add a field full_name containing the first_name and last_name of each employee separated by a space.

```{sql connection=db_connect_3}

-- This will pull the first name and last name anc combine into full name

SELECT
    first_name,
    last_name,
    CONCAT(first_name, ' ', last_name) AS full_name
FROM employees 
WHERE first_name IS NOT NULL AND last_name IS NOT NULL 
  
```

Retrieve the details of all employees again, but this time listing their team name (as team_name), rather than their team_id.

```{sql connection=db_connect_3}
--This will list all the people including their team id and their team (name) as team_name

SELECT 
  first_name,
  last_name,
  team_id,
  name AS team_name
FROM employees AS e
LEFT JOIN teams AS t
ON e.team_id = t.id
```

Bit Harder] Do a breakdown of the number of employees per team (here we want two columns: team_name and num_employees). Order the table in decreasing number of employees. Include all teams, even if they have no employees.


```{sql connection=db_connect_3}

SELECT 
  t.name AS team_name,
  COUNT(e.id) AS num_employees 
FROM teams AS t
LEFT JOIN employees AS e
ON e.team_id = t.id
GROUP BY t.name
ORDER BY num_employees DESC

```

[Bit Harder] AcmeCorp management want to send an email survey to each of their employees. However, they are worried that some of the employees may not receive the email either because their email addresses may be invalid (in that they don’t contain an ‘@’ symbol), or they may not have a stored email address at all. Can you find all these employees?



```{sql connection=db_connect_3}
-- This will bring back all the employees who dont have an at, need to get nullss also

SELECT 
  *
FROM employees
WHERE email NOT LIKE '%@%' OR email IS NULL
```


```{sql connection=db_connect_3}

```


```{sql connection=db_connect_3}

```




Run at the end to dicsonnect from the database

```{r}

dbDisconnect(db_connect_3)

```

