---
title: "R Notebook"
output: html_notebook
---

```{r}

library(RPostgreSQL)

```

```{r}

home_location <- Sys.getenv("HOME")
credentials_location <- paste0(home_location,"/credentials.R")
source(credentials_location)

db_connect_3 <- dbConnect(drv = PostgreSQL(), user = postgresql_username, password = postgresql_password, dbname = "acme_employees", host = "localhost")
rm(postgresql_username, postgresql_password)


```

Now use SQL queries within R to answer all of the following questions.

Create a new data.frame containing the details of all employees held in the database


```{sql connection=db_connect_3}

SELECT * 
FROM employees AS e
INNER JOIN Teams AS T
ON e.id = t.id

```

SOLUTION

```{r}
all_employees <- dbGetQuery(db_connect_3, "SELECT * FROM employees")
all_employees

```



Retrieve the records for all employees again, but this time add a field full_name containing the first_name and last_name of each employee separated by a space.

```{sql connection=db_connect_3}

-- This will pull the first name and last name anc combine into full name

SELECT
    first_name,
    last_name,
    CONCAT(first_name, ' ', last_name) AS full_name
FROM employees 
WHERE first_name IS NOT NULL AND last_name IS NOT NULL 
  
```

SOLUTION

```{sql connection=db_connect_3}
SELECT 
  *, 
  CONCAT(first_name, ' ', last_name) AS full_name 
FROM employees
```


Retrieve the details of all employees again, but this time listing their team name (as team_name), rather than their team_id.

```{sql connection=db_connect_3}
--This will list all the people including their team id and their team (name) as team_name

SELECT 
  first_name,
  last_name,
  team_id,
  name AS team_name
FROM employees AS e
LEFT JOIN teams AS t
ON e.team_id = t.id
```

SOLUTION

```{sql connection=db_connect_3}
SELECT 
  e.first_name, 
  e.last_name, 
  e.email, 
  t.name AS team_name
FROM employees AS e LEFT JOIN teams AS t
ON e.team_id = t.id
```


Bit Harder] Do a breakdown of the number of employees per team (here we want two columns: team_name and num_employees). Order the table in decreasing number of employees. Include all teams, even if they have no employees.


```{sql connection=db_connect_3}

SELECT 
  t.name AS team_name,
  COUNT(e.id) AS num_employees 
FROM teams AS t
LEFT JOIN employees AS e
ON e.team_id = t.id
GROUP BY t.name
ORDER BY num_employees DESC

```

SOLIUTION

```{sql connection=db_connect_3}
SELECT 
  t.name AS team_name, 
  COUNT(e.id) AS num_employees
FROM teams AS t 
LEFT JOIN employees AS e 
ON e.team_id = t.id
GROUP BY t.name
ORDER BY num_employees DESC
```


[Bit Harder] AcmeCorp management want to send an email survey to each of their employees. However, they are worried that some of the employees may not receive the email either because their email addresses may be invalid (in that they don’t contain an ‘@’ symbol), or they may not have a stored email address at all. Can you find all these employees?



```{sql connection=db_connect_3}
-- This will bring back all the employees who dont have an at, need to get nullss also

SELECT 
  *
FROM employees
WHERE email NOT LIKE '%@%' OR email IS NULL
```

SOLUTION

```{sql connection=db_connect_3}
SELECT * FROM employees
WHERE email NOT LIKE '%@%' OR email IS NULL

```

EXTENSION
2.1 HARDER SQL QUESTIONS

Which of the employees provided the most recent feedback? What was the date of this feedback?


```{sql connection=db_connect_3}

SELECT 
  e.id, 
  e.first_name, 
  e.last_name, 
  f.date 
FROM employees AS e INNER JOIN feedbacks AS f
ON f.employee_id = e.id
WHERE f.date = (SELECT MAX(date)
                  FROM feedbacks)

```

You are chatting with one of the AcmeCorp employees, and she mentions that one of the teams in the corporation has some employees whose first_name contains only their first initial, e.g. “D” instead of “Dionne”, “C” instead of “Charles” (they had a really formal administrator a few years back). Can you find the name of the team that she is most likely referring to (i.e. which team has the highest number of employees with single initial first_names)`?

```{sql connection=db_connect_3}

SELECT 
  COUNT(*) AS num_only_first_initial, 
  t.name AS team_name 
FROM employees AS e
LEFT JOIN teams AS t
ON e.team_id = t.id
WHERE e.first_name LIKE '_'
GROUP BY team_name
ORDER BY num_only_first_initial DESC NULLS LAST
LIMIT 1
```


2.2 Combining SQL and R coding - basic sentiment analysis

AcmeCorp would like you to perform very basic ‘sentiment analysis’ of the stored employee feedbacks. They have provided a set of ‘positive’ and ‘negative’ words to use to score feedback messages in this analysis.
    The score for each message should start at zero.
    Every time a ‘positive’ word occurs in a message, the message’s score should increase by 1 point.
    Every time a ‘negative’ word occurs in a message, the message’s score should decrease by 1 point.
    Punctuation marks should be removed prior to scoring [Hint - the gsub() function can help here. You probably want to pass fixed = TRUE into this function].
    Messages should also be converted to lowercase prior to scoring [Hint - the tolower() function can help here].

Write R code to score all of the stored feedback messages according to these rules.

SOLUTION

```{r}

positive_words <- c("great", "good", "happy", "content", "fulfilled", "fulfilled", "positive", "positives", "cheerful", "nice", "ambition", "ambitions", "ambitious", "friendly", "progress", "joy", "fast", "achieve", "achieving", "achievements", "best", "better",  "unlimited", "pros", "competitive", "opportunity", "advancement", "friendly", "smart", "interesting", "challenging", "grow", "caring", "balance", "helpful")

negative_words <- c("worse", "worsening", "worst", "slow", "sluggish", "impatient", "unfulfilled", "unfulfilling", "negative", "bad", "limited", "limit", "limitations", "cons", "frustrated", "frustrating", "frustration", "dull", "boring", "politics", "political", "ignore", "ignored", "ignoring", "sad", "depressed", "depressing", "nasty", "stress", "stresses", "stressed", "horrid", "horrible", "terrible", "deadline", "deadlines", "uncaring", "remote")

```


```{r}
clean_message <- function(message){
  message <- tolower(message)
  # This is messy. You'll see better ways to do this later in the course
  # We can also do something like the following, using a regular expression (regex)
  # Regexes are also covered later!
  # message <- gsub(message, pattern = "[[:punct:]]", replacement = " ")
  message <- gsub(message, pattern = ",", replacement = "", fixed = TRUE)
  message <- gsub(message, pattern = ".", replacement = "", fixed = TRUE)
  message <- gsub(message, pattern = ":", replacement = "", fixed = TRUE)
  message <- gsub(message, pattern = ";", replacement = "", fixed = TRUE)
  message <- gsub(message, pattern = "!", replacement = "", fixed = TRUE)
  message <- gsub(message, pattern = "'", replacement = "", fixed = TRUE)
  message <- gsub(message, pattern = '"', replacement = "", fixed = TRUE)
  message <- gsub(message, pattern = "(", replacement = "", fixed = TRUE)
  message <- gsub(message, pattern = ")", replacement = "", fixed = TRUE)
  message <- gsub(message, pattern = "-", replacement = " ", fixed = TRUE)
  message <- gsub(message, pattern = "/", replacement = " ", fixed = TRUE)
  message <- gsub(message, pattern = "  ", replacement = " ", fixed = TRUE)
  message <- gsub(message, pattern = "   ", replacement = " ", fixed = TRUE)
  return(message)
}

```


```{r}
score_message <- function(message, positive_words, negative_words){
  message_words <- unlist(strsplit(message, split = " ", fixed = TRUE))
  message_score <- 0
  
  for (word in positive_words){
    message_score <- message_score + sum(message_words == word)
  }
  
  for (word in negative_words){
    message_score <- message_score - sum(message_words == word)
  }
  
  return(message_score)
}
```

This is quite a tough exercise, and in reality you wouldn’t do serious sentiment analysis in this way in R. Treat this as a coding exercise; you’ll see far better approaches with dedicated text mining packages later in the course!

  First we obtained the feedbacks, with matching employees and their teams together in one dataframe.
  Next we split our solution code into two functions: clean_message() and score_message(), and ran these successively for each message stored in the database, storing the scores as we go.
  You will want to add the scores as a column in your dataframe containing feedbacks, employees and teams.

Here are examples of running both functions on two messages

```{r}
message <- "Good company, competitive salaries, nice colleagues, good location, but the food in the canteen is just terrible!"
message <- clean_message(message)
message

```


```{r}
score_message(message, positive_words, negative_words)
```

```{r}
message <- "Really cliquey, political culture, bosses ignore any complaints (and suggestions), but good benefits package"
message <- clean_message(message)
message
```

```{r}
score_message(message, positive_words, negative_words)
```

```{r}
# get dataframe with employees, teams and feedback messages
employees_teams_feedbacks <- dbGetQuery(db_connect_3,
           "SELECT e.*, f.message, t.name AS team_name
           FROM employees AS e LEFT JOIN teams AS t
           ON e.team_id = t.id INNER JOIN feedbacks AS f
           ON f.employee_id = e.id")

# go through each message in employees_teams_feedbacks, clean and score it
score <- numeric()
for (message in employees_teams_feedbacks[, 'message']){
  message <- clean_message(message)
  message_score <- score_message(message, positive_words, negative_words)
  score <- c(score, message_score)
}

# add the scores 
employees_teams_feedbacks['score'] <- score
employees_teams_feedbacks
```


One or more employees may have tied in providing the most positive comment(s). Can you find name(s) of the teams these employees belong to? Do the same for employees having provided the most negative comments. Remember that we want to do this programatically (i.e. by coding), rather than scanning through the scores by eye.
We can either do this in base R

SOLUTION

```{r}
selector <- employees_teams_feedbacks[, 'score'] == max(employees_teams_feedbacks[, 'score']) 
mostpos_employees_teams <- employees_teams_feedbacks[selector, "team_name"]
mostpos_employees_teams

```

```{r}
selector <- employees_teams_feedbacks[, 'score'] == min(employees_teams_feedbacks[, 'score']) 
mostneg_employees_teams <- employees_teams_feedbacks[selector, "team_name"]
mostneg_employees_teams
```

or, alternatively, we can use sqldf

```{r}
install.packages("sqldf")

```

```{r}
detach("package:RPostgreSQL", unload = TRUE)
library(sqldf)
```

```{r}
teams_mostpos_employees_querystring <- "SELECT team_name
                                        FROM employees_teams_feedbacks
                                        WHERE score = (SELECT MAX(score)
                                        FROM employees_teams_feedbacks)"

mostpos_employees_teams <- sqldf(teams_mostpos_employees_querystring)
mostpos_employees_teams
```

```{r}
teams_mostneg_employees_querystring <- "SELECT team_name
                                        FROM employees_teams_feedbacks
                                        WHERE SCORE = (SELECT MIN(score)
                                        FROM employees_teams_feedbacks)"

mostneg_employees_teams <- sqldf(teams_mostneg_employees_querystring)
mostneg_employees_teams
```


Run at the end to dicsonnect from the database

```{r}

dbDisconnect(db_connect_3)

```


