---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
```


```{r}

house_data <- read_csv("data/kc_house_data.csv")

```

CODECLAN SOILUTION

```{r}
houses <- read_csv("data/kc_house_data.csv")
```

```{r}
glimpse(houses)
```


```{r}
library(modelr)
```



```{r}
house_data
```


Tidy up the data ready for regression:

You might like to think about removing some or all of date, id, sqft_living15, sqft_lot15 and zipcode (lat and long provide a better measure of location in any event).

```{r}
house_data_tidy <- house_data %>%
  select(-c("date", "id", "sqft_living15", "sqft_lot15", "zipcode"))

house_data_tidy
```

CODECLAN SOLUTION

```{r}
# tidy up data. In particular treat condition and grade as factor, as they are
# ordinal categorical
houses_tidy <- houses %>%
  select(-c("id", "date", "sqft_living15", "sqft_lot15", "zipcode")) %>%
  mutate(waterfront = as.logical(waterfront)) %>%
  mutate(renovated = yr_renovated != 0) %>%
  select(-"yr_renovated") %>%
  mutate(condition = as_factor(condition)) %>%
  mutate(grade = as_factor(grade))

glimpse(houses_tidy)
```


Have a think about how to treat waterfront. Should we convert its type?

We converted yr_renovated into a renovated logical variable, indicating whether the property had ever been renovated. You may wish to do the same.

Have a think about how to treat condition and grade? Are they interval or categorical ordinal data types?


Check for aliased variables using the alias() function (this takes in a formula object and a data set). [Hint - formula price ~ . says ‘price varying with all predictors’, this is a suitable input to alias()]. Remove variables that lead to an alias. Check the ‘Elements of multiple regression’ lesson for a dropdown containing further information on finding aliased variables in a dataset.

```{r}
library(modelr)
```


```{r}
alias(lm(price~ ., data = house_data_tidy))
```

```{r}
house_price_tidy <- house_data_tidy %>%
  select(-c("sqft_living", "sqft_above"))
```

CODECLAN SOLUTION

```{r}
# Alias is useful to check if we have aliased variables, i.e. one or more
# variables that can be computed from other variables
alias(price ~ ., data = houses_tidy)

```

```{r}
# seems that sqft_basement can be computed from sqft_living - sqft_above.
# let's drop sqft_living leaving just the two contributions sqft_basement and 
# sqft_above
houses_tidy <- houses_tidy %>%
  select(-"sqft_living")

glimpse(houses_tidy)
```


Systematically build a regression model containing up to four main effects (remember, a main effect is just a single predictor with coefficient), testing the regression diagnostics as you go

splitting datasets into numeric and non-numeric columns might help ggpairs() run in manageable time, although you will need to add either a price or resid column to the non-numeric dataframe in order to see its correlations with the non-numeric predictors.

```{r}
library(GGally)
```


```{r}
houses_tidy_numeric <- house_price_tidy %>%
  select_if(is.numeric)

houses_tidy_nonnumeric <- house_price_tidy %>%
  select_if(function(x) !is.numeric(x))

houses_tidy_nonnumeric$price <- house_price_tidy$price

ggpairs(houses_tidy_numeric)
ggpairs(houses_tidy_nonnumeric)
```


```{r}
mod1a <- lm(price ~ grade, data = houses_tidy_numeric)

mod1a
```


```{r}
summary(mod1a)
```

```{r}
par(mfrow = c(2,2))
plot(mod1a)
```


```{r}
mod2a <- lm(price ~ condition, data = house_data_tidy)

mod2a
```


```{r}
summary(mod2a)
```

```{r}
house_remianing_resid <- houses_tidy_numeric %>%
  add_residuals(mod1a) %>%
  select(-c("price", "grade"))

house_remianing_resid %>%
  ggpairs(aes(colour = condition, alpha = 0.5))
```


```{r}

par(mfrow = c(2,2))
plot(mod2a)

```

```{r}
anova(mod1a, mod2a)
```

```{r}

mod1b <- lm(price ~ long, data = houses_tidy_numeric)

mod1b

```


```{r}
summary(mod1b)
```

CODECLAN SOLUTION

```{r}
houses_tidy_numeric <- houses_tidy %>%
  select_if(is.numeric)

houses_tidy_nonnumeric <- houses_tidy %>%
  select_if(function(x) !is.numeric(x))

houses_tidy_nonnumeric$price <- houses_tidy$price

ggpairs(houses_tidy_numeric)

```

```{r}
ggpairs(houses_tidy_nonnumeric)
```

Correlation of sqft_above with price looks pretty promising, but split of price by grade and waterfront also look decent.

```{r}
houses_tidy %>%
  ggplot(aes(x = grade, y = price)) +
  geom_boxplot()
```

```{r}
houses_tidy %>%
  ggplot(aes(x = waterfront, y = price)) +
  geom_boxplot()
```

```{r}
mod1_a <- lm(price ~ sqft_above, data = houses_tidy)
summary(mod1_a)
```

```{r}
mod1_b <- lm(price ~ grade, data = houses_tidy)
summary(mod1_b)
```

```{r}
mod1_c <- lm(price ~ waterfront, data = houses_tidy)
summary(mod1_c)
```

```{r}
# grade looks the most promising, but some of the grade level coeffs are insignificant.
# the F-test at the bottom of the regression output tests against the null model (i.e. intercept only)
# but, if we want, we can replicate this using a separate anova
# null model: regress price on intercept only
null_model <- lm(price ~ 1, data = houses_tidy)
grade_model <- lm(price ~ grade, data = houses_tidy)
anova(null_model, grade_model)
```

```{r}
# grade is significant, let's keep it. Now plot diagnostics
par(mfrow = c(2, 2))
plot(mod1_b)
```

```{r}
houses_resid <- houses_tidy %>%
  add_residuals(mod1b) %>%
  select(-c("price", "grade"))

houses_resid_numeric <- houses_resid %>%
  select_if(is.numeric)

houses_resid_nonnumeric <- houses_resid %>%
  select_if(function(x) !is.numeric(x))

houses_resid_nonnumeric$resid <- houses_resid$resid
```

```{r}
ggpairs(houses_resid_numeric)
```


```{r}
ggpairs(houses_resid_nonnumeric)
```

lat has highest correlation with residuals, but, again, waterfront still looks pretty promising. Try both…

```{r}
mod2_a <- lm(price ~ grade + lat, data = houses_tidy)
summary(mod2_a)
```

```{r}
mod2_b <- lm(price ~ grade + waterfront, data = houses_tidy)
summary(mod2_b)
```

```{r}
# lat is significant and higher r^2, let's keep model2a
par(mfrow = c(2, 2))
plot(mod2_a)
```

```{r}
houses_resid <- houses_tidy %>%
  add_residuals(mod2_a) %>%
  select(-c("price", "grade", "lat"))

houses_resid_numeric <- houses_resid %>%
  select_if(is.numeric)

houses_resid_nonnumeric <- houses_resid %>%
  select_if(function(x) !is.numeric(x))

houses_resid_nonnumeric$resid <- houses_resid$resid
```

```{r}
ggpairs(houses_resid_numeric)
```

```{r}
ggpairs(houses_resid_nonnumeric)
```

Now view has strongest correlation with residuals, but also compare against model with waterfront.

```{r}
mod3_a <- lm(price ~ grade + lat + view, data = houses_tidy)
summary(mod3_a)
```

```{r}
mod3_b <- lm(price ~ grade + lat + waterfront, data = houses_tidy)
summary(mod3_b)
```

```{r}
# view model is best, keep mod3a
par(mfrow = c(2, 2))
plot(mod3_a)
```

```{r}
houses_resid <- houses_tidy %>%
  add_residuals(mod3_a) %>%
  select(-c("price", "grade", "lat", "view"))

houses_resid_numeric <- houses_resid %>%
  select_if(is.numeric)

houses_resid_nonnumeric <- houses_resid %>%
  select_if(function(x) !is.numeric(x))

houses_resid_nonnumeric$resid <- houses_resid$resid
```

```{r}
ggpairs(houses_resid_numeric)
```


```{r}
ggpairs(houses_resid_nonnumeric)
```

sqft_basement has highest correlation with residuals. Let’s test against all remaining categorical predictors:

```{r}
mod4_a <- lm(price ~ grade + lat + view + sqft_basement, data = houses_tidy)
summary(mod4_a)
```

```{r}
mod4_b <- lm(price ~ grade + lat + view + waterfront, data = houses_tidy)
summary(mod4_b)
```

```{r}
mod4_c <- lm(price ~ grade + lat + view + condition, data = houses_tidy)
summary(mod4_c)
```

```{r}
mod4_d <- lm(price ~ grade + lat + view + renovated, data = houses_tidy)
summary(mod4_d)
```

```{r}
# looks like model with sqft_basement is best, keep mod4a
par(mfrow = c(2, 2))
plot(mod4_a)
```

```{r}
houses_resid <- houses_tidy %>%
  add_residuals(mod4_a) %>%
  select(- price)
```

Our final model in terms of main effects is: price ~ grade + lat + view + sqft_basement

EXTENSION

Consider possible interactions between your four main effect predictors and test their effect upon r2. Choose your best candidate interaction and visualise its effect.

Calculate the relative importance of predictors from your best 4-predictor model (i.e. the model without an interaction). Which predictor affects price most strongly?

Now, for interactions, have six possibilities that obey principle of strong hierarchy (i.e. consider including an interaction only if its main effects are already present in the model)

```{r}
mod5_a <- lm(price ~ grade + lat + view + sqft_basement + grade:lat, data = houses_tidy)
summary(mod5_a)
```

```{r}
mod5_b <- lm(price ~ grade + lat + view + sqft_basement + grade:view, data = houses_tidy)
summary(mod5_b)
```

```{r}
mod5_c <- lm(price ~ grade + lat + view + sqft_basement + grade:sqft_basement, data = houses_tidy)
summary(mod5_c)
```

```{r}
mod5_d <- lm(price ~ grade + lat + view + sqft_basement + lat:view, data = houses_tidy)
summary(mod5_d)
```

```{r}
mod5_e <- lm(price ~ grade + lat + view + sqft_basement + lat:sqft_basement, data = houses_tidy)
summary(mod5_e)
```

```{r}
mod5_f <- lm(price ~ grade + lat + view + sqft_basement + view:sqft_basement, data = houses_tidy)
summary(mod5_f)
```

```{r}
# mod5c looks like the best
par(mfrow = c(2,2))
plot(mod5_c)
```

It seems that the grade:sqft_basement interaction leads to highest r2 (but two levels of the interaction cannot be determined due to fitting problems).

Now let’s see a visualisation of the effect of this interaction.

```{r}
houses_resid %>%
  ggplot(aes(x = sqft_basement, y = resid, colour = grade)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE) +
  facet_wrap(~ grade)
```

Relative importance of predictors:

```{r}
library(relaimpo)
```

```{r}
calc.relimp(mod4_a, method = "lmg", rela = TRUE)
```

It looks like the grade of property is the most important determiner of price, followed by the number of views the property has received.

