---
title: "R Notebook"
output: html_notebook
---

```{r}

library(RPostgreSQL)

```

```{r}
home_location <- Sys.getenv("HOME")
credentials_location <- paste0(home_location,"/credentials.R")
source(credentials_location)

db_connect <- dbConnect(drv = PostgreSQL(), user = postgresql_username, password = postgresql_password, dbname = "omni_employees", host = "localhost")
rm(postgresql_username, postgresql_password)

```

Find all the employees who work in the ‘Human Resources’ department.

```{sql connection=db_connect}
SELECT *
FROM employees
WHERE department = 'Human Resources'

-- this iwll return list of mployees in HE department
```

Get the first name, last name, and country of the employees who work in the ‘Legal’ department


```{sql connection=db_connect}

SELECT
  first_name,
  last_name, 
  country
FROM employees
WHERE department = 'Legal'

-- this iwll return list of ee's first & last name and country in the legal dept

```

Count the number of employees based in Portugal.


```{sql connection=db_connect}

-- this iwll return count of employees in portugal

SELECT 
  COUNT(*) AS number_employees_in_Portugal
FROM employees
WHERE country = 'Portugal'

```

Count the number of employees based in either Portugal or Spain.

```{sql connection=db_connect}
-- this will pull the bumber of employees in Portugal

SELECT 
  COUNT(id) 
FROM employees 
WHERE country IN ('Spain', 'Portugal')

```

Count the number of pay_details records lacking a local_account_no 


```{sql connection=db_connect}
-- this bring the number of employees with no account no
SELECT 
  COUNT(*) AS employee_account  
FROM pay_details
WHERE local_account_no IS NULL

```

Are there any pay details record lacking both account 

```{sql connection=db_connect}

SELECT 
  COUNT(id)
FROM pay_details
WHERE local_account_no is NULL

```


```{r}

dbListTables(conn = db_connect)

```


Are there any pay_details record lacking both a local_account_no and iban number?


```{sql connection=db_connect}

-- This will show the N/A's on the local account, still need to pull iban too

SELECT 
  COUNT(id)
FROM pay_details
WHERE local_account_no iS NULL and Iban IS NULL

```


Get a list of employees first names and last names ordered alphabetically by their last name (put any nulls last).

```{sql connection=db_connect}

--this bring back the employees first and last nanmes aphabetically

SELECT 
  first_name,
  last_name
FROM employees
ORDER BY last_name ASC NULLS LAST

```

Get a list of employees first names, last names and countries, ordered alphabetically first by their country and second by their last name (put any nulls last).


```{sql connection=db_connect}

-- this will show employees ordered alphabetically by country first

SELECT 
  first_name,
  last_name,
  country
FROM employees
ORDER BY country, last_name ASC NULLS LAST

```

Find the top ten highest paid employees in the corporation.

```{sql connection=db_connect}

SELECT 
  Id,
  first_name,
  last_name, 
  salary
FROM employees
ORDER BY salary DESC NULLS LAST 
LIMIT 10


```

Find the first name, last name and salary of the lowest paid of the employees in Hungary


```{sql connection=db_connect}

SELECT
  first_name,
  last_name, 
  salary, 
  country,
FROM employees
WHERE country = 'Hungary'
ORDER BY salary ASC NULLS LAST 


```

How many employees have a first name beginning with ‘F’?

```{sql connection=db_connect}

-- This brings back all employees with surname startying f

SELECT 
  COUNT(id)
FROM employees
WHERE first_name LIKE '%F%'

```

Find all the details of any employees with a ‘yahoo’ e-mail address?

```{sql connection=db_connect}

SELECT *
FROM employees
WHERE email LIKE ('%yahoo%')

```



Count the number of pension enrolled employees not based in either France or Germany.


```{sql connection=db_connect}

SELECT
  COUNT(id) AS pens_enrolled
FROM employees 
WHERE pension_enrol = TRUE AND country NOT IN ('France', 'Germany')
```

Breakdown the numbers of employees enrolled, not enrolled and whose enrollment status is unknown in the corporation pension scheme.

```{sql connection=db_connect}

SELECT 
  pension_enrol,
  COUNT (id) AS num_employees
FROM employees
GROUP BY pension_enrol

```

What is the maximum salary among employees in the Engineering department who work 1.0 full-time equivalent hours?

```{sql connection=db_connect}
-- max execution not working and not sure why

SELECT 
  MAX (salary) AS max_salary
FROM employees 
WHERE department = 'Engineering' AND fte_hours = 1.0
```

Obtain a count by department of the employees who started work with the corporation in 2003.

```{sql connection=db_connect}

SELECT 
  department,
  COUNT(id) AS number_of_emp
FROM employees
WHERE start_date BETWEEN '2003-01-01' AND '2003-12-31'
GROUP BY department

```



```{sql connection=db_connect}



```



Do this at the end


```{r}
dbDisconnect(db_connect)
```

