---
title: "R Notebook"
output: html_notebook
---

```{r}
library(readr)
```


1. Read in the state_income_data.csv file usine the approprial function


```{r}  

## This reads in the csv file using

income_data <- read_csv("data/state_income_data.csv")

```



```{r}
## imported data

income_data
```


```{r}

class(income_data)

View(income_data)
```


2. How do you specify your own column names when using read_csv?

```{r}

## thisa is an example of the code that would be used to speicy own column names
 
income_data <- read.csv("data/state_income_data.csv", skip = 1, 
                        col_names = c( "Index",
                                       "state", 
                                       "y2002"
                                      ))           
                                

```

SOLUTION

```{r}
# 2. How do you specify your own column names when using `read_csv`?
col_names =

```


3. read in cars_csv file and use the col_types() function to ensure you load in x1 as char, mpg, cyl, disp, hp as double and vs as logical.


```{r}
## Loaded in thge csv file

cars_data <- read_csv("data/cars.csv")

```

```{r}

cars_data

```


```{r}
class(cars_data)

```



```{r}
## code input to change the csv file, so the vs column became logical

cars_data <- read_csv("data/cars.csv", col_types = cols(X1 = col_logical()))

```


SOLUTION

```{r}
# 3. Read in the `cars.csv` file and use the 'col_types' function to ensure you load in:
# `X1` as character, `mpg`, `cyl`, `disp`, `hp` as double, `vs` as logical
car_data <- read_csv("data/cars.csv", col_types = cols( X1 = col_character(), 
                     mpg = col_double(),
                     cyl = col_double(),
                     disp = col_double(),
                     hp = col_double(),
                     vs = col_logical()))



```


4. read the Scotland.txt file using the appopriate function


```{r}

## This is the function to read text file

scotland_text <- read_file("data/scotland.txt")

class(scotland_text)

scotland_text
```

SOLUTION

```{r}
scotland <- read_delim("data/scotland.txt", delim = ";")
```


5. How do you read text into one element character vector?

```{r}
## this read the text as one element

scotland_two <- read_file("data/scotland.txt")

scotland_two
```

SOLUTION

```{r}
# 5. How do you read text into a one element character vector?
read_file
```


6. Read in the steam_200k_sample.csv file. fix any problems you may encounter



```{r}

## when loading the code this error is received due to errors in the first 2 rows of the data

steam_200k_data <- read.csv("data/steam-200k-sample.csv")

```


```{r}

## This will skip the first two rows when it reads the data in

steam_200k_data <- read.csv("data/steam-200k-sample.csv", skip = 2)

steam_200k_data
```

SOLUTION

```{r}
# 6. Read in the `steam-200k-sample.csv` file. Fix any problems you may encounter. 
# load in the video games data, and remove the metadata 
video_games <- read_csv("data/steam-200k-sample.csv",skip = 4)

```

7. read in the historic_eu_payments.tsv file. 

```{r}
## This is how you read tsv file in

historic <- read_tsv("data/historic_eu_payments.tsv")
```

SOLUTION

```{r}
# 7. Read in the `historic_eu_payments.tsv` file. 
eu_payments <- read_tsv("data/historic_eu_payments.tsv")

```

SOLUTION

```{r}
# 7. Read in the `historic_eu_payments.tsv` file. 
eu_payments <- read_tsv("data/historic_eu_payments.tsv")


```

8. How do you read lines of text into a multi element character vector

```{r}

## this is to reead line of text into multi elemnt character vector
  ## this is done usong the delim code e.g

beer <- read_delim("data/beer.txt", delim = ";")

```

SOLUTION

```{r}
# 8. How do you read lines of text into a multi-element text vector?
read_lines

```


9. Read in the JSON file from the following url: http://makemeanalyst.com/wp-content/uploads/2017/06/sample_json.txt


```{r}

library(jsonlite)
library(readr)

```



```{r}
## This is how to read in a JSON file

raw_makeme <- read_json("http://makemeanalyst.com/wp-content/uploads/2017/06/sample_json.txt")

```

```{r}

## This is the Json file
raw_makeme

```

```{r}

class(raw_makeme)
```

SOLTUTION

```{r}
raw_json <-  read_file("http://makemeanalyst.com/wp-content/uploads/2017/06/sample_json.txt")
parse_json(raw_json)
```

EXTENSION

A lot of Excel sheets you will encounter will contain more than one sheet. Try to write some code that reads in all the different sheets in the file green_bindle_consulting.xls

SOLUTION

```{r}
library(readxl)

```

```{r}

## Method 1 (no loops) - lappy & naming the sheets the names of the tabs.
name_sheets<-excel_sheets("data/green_bindle_consulting.xlsx")
num_sheets<-length(name_sheets)

all_tabs <- lapply(name_sheets, function(x) read_excel("data/green_bindle_consulting.xlsx", sheet = x))
names(all_tabs)<-name_sheets
#list2env(all_tabs, globalenv()) #may not be necessary to do this step - may want to just call each element of the list

```


```{r}
## Method 2  - start with a list and loop through assigning elements of the list
all_tabs <- list()
name_sheets<-excel_sheets("data/green_bindle_consulting.xlsx")

for(sheet in name_sheets) {
  all_tabs[[sheet]] <- data.frame(read_excel("data/green_bindle_consulting.xlsx", 
                                                    sheet = sheet))
}

```


```{r}
## Method 3 - loops & naming the sheets 'data_sheet_1' etc.
num_sheets<-length(excel_sheets("data/green_bindle_consulting.xlsx"))

for(i in 1:num_sheets) {
  assign(paste0("data_sheet_",i), data.frame(read_excel("data/green_bindle_consulting.xlsx", 
                                                sheet = i)))
}
```


```{r}
## Method 4 - loops & naming the sheets the names of the tabs.
name_sheets<-excel_sheets("data/green_bindle_consulting.xlsx")
num_sheets<-length(name_sheets)

for(i in 1:num_sheets) {
  assign(name_sheets[i], data.frame(read_excel("data/green_bindle_consulting.xlsx", 
                                                        sheet = i)))
  }

```

