---
title: "R Notebook"
output: html_notebook
---

code to install and load dataset packages

```{r}
install.packages("devtools")
```

```{r}
devtools::install_github('codeclan/CodeClanData')

```

```{r}
library(CodeClanData)
```

```{r}
beer
```


```{r}
class(beer)
```

```{r}
str(beer)
```

The variable calories is character type and should be numeric.

SOLUTION

```{r}
beer$calories <- as.numeric(beer$calories)
```


what are the mean values of all the numeric variables
```{r}
beer[ ,3:5]
#this shows the last 3 columns.
```


```{r}
max(beer[ ,3])

#maximum of percentage column
```

```{r}
min(beer[ ,3])

#minimum of percentage column
```

```{r}
mean(beer[ ,3])
#average percentage
```

```{r}
mean(beer[ ,5])
#avenage carb
```


SOLUTION

```{r}
mean(beer$percent)
```

```{r}

mean(beer$calories)

```


```{r}

mean(beer$carbohydrates)
```


Do the mean calories for high, medium and low alcohol beers differ significantly?

```{r}
low_alcohol <- subset(beer, percent < 2)
medium_alcohol <- subset(beer, percent >= 2 & percent < 5)
high_alcohol <- subset(beer, percent >= 5)

mean(low_alcohol$calories)

```

```{r}
mean(medium_alcohol$calories)
```

```{r}
mean(high_alcohol$calories)
```

Which beer has the highest percent?

```{r}
subset(beer, percent == max(percent))
```

Which beer has the highest calories?

```{r}
subset(beer, calories == max(calories))
```


```{r}
temp_data_b$brand
#this is all the bramds showing
```

```{r}
temp_data_b[ , 1:2]
```

```{r}
temp_data_b[ ,-4]
#everything except the calories
```

```{r}
temp_data_b[ 3 < 6]
```


```{r}
sum(temp_data_t)
#adds up all elements 
```

```{r}
mean(temp_data_t)

#average of temp
```

```{r}
str(temp_data_t)
```
1.2 TEMPERATURE

SOLUTION

What was the average maximum temperature for each month?

```{r}
colMeans(temp)
```


What was the average maximum monthly temperature for each year?

```{r}
rowMeans(temp)
```

What was the highest maximum monthly temperature recorded?

```{r}
max(temp)
```

In how many months was the maximum recorded temperature greater than 20?

```{r}
sum(temp > 20)
```

```{r}
# or
length(temp[temp > 20])
```

1.3 GAME OF THRONES

```{r}
game_of_thrones
```



```{r}
temp_data_b <- CodeClanData::beer
```


```{r}
temp_data_t <- CodeClanData::temp
```

```{r}
str(temp_data)
#this shows the structure of the gane of thrones package
```

```{r}
str(temp_data_b)
#this shows the structure of the beed df
```

```{r}
str(temp_data_t)
#the structre of the temperature dataset
```

SOLUTION

How many characters are there in the book?

```{r}
length(game_of_thrones$characters)

```

How many pages are there in the book?

```{r}
game_of_thrones$numberOfPages
```

What are the full details for Robb Stark

```{r}
game_of_thrones$characters[[397]]

```

2. EXTENSION

2.1 BEER

Plot how calories varies with percent

```{r}
plot(beer$percent, beer$calories)
```

Plot how calories varies with carbohydrates

```{r}
plot(beer$percent, beer$carbohydrates)
```

[Harder] Summarise the beers produced by each brewery.

```{r}
numeric_beer <- beer[, 3:5]
numeric_beer <- na.omit(numeric_beer)
by(numeric_beer, beer$brewer, summary)

```

2.2 TEMPERATURE

Plot the average maximum temperature for each year

```{r}
plot(rowMeans(temp))

```


Plot a histogram showing the spread of temperatures

```{r}
hist(temp)
```

2.3 GAME OF THRONES

[Harder] Make a bar plot showing the distribution of gender for all characters.

```{r}
# unlisting using base R tools is pretty annoying, but we can do something like this
chars <- game_of_thrones$characters
chars_df <- as.data.frame(matrix(unlist(chars), nrow = length(chars), byrow = TRUE))
names(chars_df) <- c("Name", "Gender")

barplot(table(chars_df$Gender))
```

