---
title: "R Notebook"
output: html_notebook
---

1 MVP
Now we’ll go back to CI creation in the normal fashion. We’ll take the ames data from the CIs lab earlier today and regard it now as a sample, we won’t be drawing any smaller samples from within it. This is the usual situation in an analysis: you use all the data available to you!


Load the data again, clean_names(), and re-familiarise yourself with it

```{r}
library(tidyverse)
library(janitor)
library(infer)

```


```{r}
ames_resample <- read_csv("~/codeclan_work/codeclan_homework_DebbieL/week_06/day_03/data/ames.csv")
```


```{r}
ames_clean <- clean_names(ames_resample)

ames_clean
```


Investigate the distribution of lot_area. Is the distribution roughly normal? If not, what problems do you find?


```{r}
ames_clean_mean <- ames_clean %>%
  summarise(
    mean_lot_area = mean(lot_area))

ames_clean_mean


```
other variablle of interest

```{r}
ames_clean %>%
  ggplot(aes(x = sale_price)) +
  geom_histogram(col = "white", fill = "steel blue", alpha = 0.5)

```



Compute and visualise a bootstrap sampling distribution for the mean(lot_area) of the sold houses.



```{r}
bootstrap_resample_200 <- ames_clean %>%
                      rep_sample_n(size = 200, replace = TRUE, reps = 200) %>%
                      summarise(
                      mean_lot_area = mean(lot_area)
                      )

bootstrap_resample_200
```


```{r}
bootstrap_resample_200 %>%
  ggplot(aes(x = mean_lot_area)) +
  geom_histogram(col = "white", fill = "steel blue", alpha = 0.7)

```


Use your bootstrap distribution to calculate a 95% CI for mean(lot_area), and visualise it on the distribution


```{r}
ci_95 <- bootstrap_resample_200 %>%
  summarise(
    mean = mean(mean_lot_area),
    lower_bound = quantile(mean_lot_area, probs = 0.025),
    upper_bound = quantile(mean_lot_area, probs = 0.975)
  )

ci_95

```

Bootstrap distribution

```{r}
infer_resample <- ames_clean %>%
      specify(response = lot_area) %>%
      generate(reps = 200, type = "bootstrap") %>%
      calculate(stat = "mean")

infer_resample

```


```{r}
infer_resample_95 <- infer_resample %>%
        get_confidence_interval(level = 0.95, type = "percentile")

infer_resample_95

```

```{r}
infer_resample %>%
  visualise(bins = 30) +
  shade_confidence_interval(endpoints = infer_resample_95)

```


You would like to know the mean(lot_area) of the sold houses with higher confidence. Calculate the 99% CI for this variable (you can re-use your bootstrap distribution from above). Is it narrower or broader than the 95% CI? Does that make sense?


```{r}

ci_99 <- bootstrap_resample_200 %>%
  summarise(
    mean = mean(mean_lot_area),
    lower_bound = quantile(mean_lot_area, probs = 0.005),
    upper_bound = quantile(mean_lot_area, probs = 0.995)
  )

ci_99
```



Calculate the point estimate of the mean(lot_area)

```{r}
infer_resample %>%
  summarise(mean(stat))

```


2 Extension


Calculate a point estimate and 95% CI for the proportion of houses in the data built before 1920. Does the number of reps you use matter? [Investigate reps from 200 up to 50000, memory of your laptop permitting].

[Hint - the current implementation of calculate(stat = "prop") in infer is slow! You can get around this by treating the mean in this way: add a new column via mutate(before_1920 = as.numeric(year_built < 1920)) and then calculate(stat = "mean") on this new column]


```{r}
ames_built <- ames_clean %>%
      filter(year_built < 1920)

ames_built

```

```{r}

bootstrap_ames_built_176 <- ames_built %>%
    rep_sample_n(size = 176, replace = TRUE, reps = 176) %>%
    summarise(
      mean_year_built = mean(year_built)
    )

bootstrap_ames_built_176
  
```


```{r}

bootstrap_ames_built_176 %>%
  ggplot(aes(x = mean_year_built)) +
  geom_histogram(col = "white", fill = "steel blue", alpha = 0.7)

```


```{r}
ames_ci_95 <- bootstrap_ames_built_176 %>%
  summarise(
    mean = mean(mean_year_built),
    lower_bound = quantile(mean_year_built, probs = 0.025),
    upper_bound = quantile(mean_year_built, probs = 0.975)
  )

ames_ci_95

```


```{r}
infer_ames_built <- ames_built %>%
  specify(response = year_built) %>%
  generate(reps = 176, type = "bootstrap") %>%
  calculate(stat = "mean")

infer_ames_built

```


```{r}
infer_ci_95_ames <- infer_ames_built %>%
  get_confidence_interval(level = 0.95, type = "percentile")

infer_ci_95_ames

```

```{r}
infer_ames_built %>%
  visualise(bins = 30) +
  shade_confidence_interval(endpoints = infer_ci_95_ames)

```


```{r}
infer_ames_built %>%
  summarise(mean(stat))

```



```{r}
ames_built_2 <- ames_built %>%
  mutate(before_1920 = as.numeric(year_built < 1920))

ames_built_2

```

```{r}
bootstrap_ames_built2_176 <- ames_built_2 %>%
    rep_sample_n(size = 176, replace = TRUE, reps = 176) %>%
    summarise(
      mean_year_built = mean(before_1920)
    )


bootstrap_ames_built2_176
```

