---
title: "R Notebook"
output: html_notebook
---


2.2.1 MULTIPLY number by 15 and add 4 to give output

```{r}
## Set z to number 4 and y to number 3 and wrote if code that states if y is greater than or equal to 0, then the number will be multiplied by 15 and then 4 will bw added on

z <- 4

y <- 1

if (y >= 0) {
  y * 15 + z
}
    


```

SOLUTION

Write a function which multiples all numbers by 15 and adds 4.

```{r}
times_15_add_4 <- function(x) {
  x * 15 + 4
}
```

2.2.1 Write a function tht returns "bird, if the input is goose, duck or pigeon and returns not bird for everythign else


```{r}

my_animals <- c("goose", "duck", "Pigeon")

## created a vector for the equired animals that are to return the bird statement

if("goose" %in% my_animals){
  print("bird") 
  ## if goose is input or any of the animals in the vector above this condition will be met and bird will be printed
} else {
    print("not a bird")
  ## anything else outside the vector wil meed this condition and return not a bird
}

```


SOLUTION

```{r}
is_bird <- function(x) {
  if (x %in% c("goose", "duck", "pigeon")) {
    return("bird")
  } else {
    return("not bird")
  }
}

```


2.2.3 write a function that muliplies all numbers by 10, if the input isnt a number the functiton should return "invalid input"

## there will be a loop needed.


```{r}
## this iwll multiply and number greater than 0 by 10, if 0 or any negative number is input this will pull message

y <- 0


if (y > 0) {
  y * 10
} else
  print("invalid input")

```

SOLUTION

```{r}
times_10 <- function(x) {
  if (is.numeric(x)) {
    return(x * 10)
  } else {
    return("Invalid input")
  }
}
```



```{r}

library(dplyr)

```


2.3.1 for the vector below, make a new vector which is equal to "big" everytime the element of sizes is greater than 10 and equal to "small" everytime the element of sizes is less than or equal to 10

```{r}

## this turns the numerical vector into a logical one

sizes <- c(13, 15, 10, 8, 22) 

sizes <= 10

```

## this code has brought back the ones that are big and any lower than or equal to 10 are small

```{r}

ifelse(sizes <= 10, "Small", "Big")

```

SOLUTION

```{r}
sizes <- c(13, 15, 10, 8, 22)

# ifelse
ifelse(sizes > 10, "Big", "Small")

```

RECODE

```{r}
# recode
recode(sizes, `10` = "Small", `8` = "Small", .default = "Big")

```

CASE WHEN

```{r}
# case_when
case_when(
  sizes > 10 ~ "Big",
  TRUE ~ "Small"
)
```

2.3.2 For the vector below convert: apple - alpaca, banana- bonobo, cherry - chipmunk

```{r}
## this recode funtion has amended the parts of the vectors to alternate words

my_fruit <- as.factor(c("apple", "apple", "banana", "cherry", "cherry"))

dplyr::recode(my_fruit, "apple" = "alpaca", "banana" = "bonono", "cherry" = "chipmunk")

```

SOLUTION

```{r}
fruit <- c("apple", "apple", "banana", "cherry", "cherry")

# if else
animals <- ifelse(fruit == "apple", "alpaca", fruit)
animals <- ifelse(animals == "banana", "bonobo", animals)
ifelse(animals == "cherry", "chipmunk", animals)

```


```{r}
# recode
recode(fruit, apple = "alpaca", banana = "bonobo", cherry = "chipmunk")
```

```{r}
# case_when
case_when(
  fruit == "apple" ~ "alpaca",
  fruit == "banana" ~ "bonobo",
  fruit == "cherry" ~ "chipmunk"
)
```


2.3.3 the two vectors below show the ages at marriage of each member of a couple, the first element  of each vector corresponds to the first couple, the second elements to the second couple 

creat a new vector which is equal to 

"Wife older" whenever the wifes ages is the highest. "Husband older" when the husbands age is the highest, "Same age" when they are the same age


```{r}
## this is the vecots creasted for husband age and wife age

wife_age <- c(29L, 30L, 36L, 31L, 22L)

husband_age <- c(28L, 35L, 32L, 31L, 29L)

```


```{r}

case_when(
  wife_age > husband_age ~ "Wife oldest",
  ## this brings back the wife oldest at the oldest point
  wife_age < husband_age ~ "husband oldest",
  ## this returns the husband oldest at oldest ages
  wife_age == husband_age  ~ "same", 
  ## this gives the same
)
  
```

SOLUTION

```{r}
wife_age <- c(29L, 30L, 36L, 31L, 22L)
husband_age <- c(28L, 35L, 32L, 31L, 29L)
```


```{r}
# ifelse
ifelse(wife_age > husband_age, "Wife older", ifelse(wife_age < husband_age, "Husband older", "Same age"))
```

```{r}

# recode
difference <- sign(wife_age - husband_age)
recode(difference, `1` = "Wife older", `0` = "Same age", `-1` = "Husband older")
```

```{r}
# case_when
case_when(
  wife_age > husband_age ~ "Wife older",
  wife_age < husband_age ~ "Husband older",
  wife_age == husband_age ~ "Same age"
)

```


2.4.1 Use a loop to print out every element of my_ist  which is type character 


```{r}

## created an apple as character function to pull the characte part of the list

apple <- as.character("apple")

my_list <- list(
  4,
  "apple",
  10.5,
  c("is", "an", "apple"),
  "apple",
  c("i", "like", "apples"),
  8
)


```


```{r}

for (elem in my_list) {
  if (is.character(elem)) {
    print(elem)
  }
}

```



```{r}

for (apple in my_list) {
  phrase <- paste(apple, "is an apple",  "i like apples", sep = " ")
print(phrase)
}
```



```{r}

length_list <- length(my_list)

```

SOLUTION

```{r}

for (elem in my_list) {
  if (is.character(elem)) {
    if (length(elem) > 1) {
      print(paste(elem, collapse = " "))
    } else {
      print(elem)
    }
  }
}

```


```{r}

##vector created for the for loop


for (i in length_list) {
  phrase <- paste("is an ", "i like apples", length_list [i])
  print(phrase)
}

##the for loop allow the code to join the vectors together by length

```


```{r}

for(i in 1:length(my_list)){
  if(class(my_list[[i]]) == "character"){
    if(length(my_list[[i]]) > 1){
      pasted_obj <- paste(my_list[[i]], collapse = " ")
      }
    return(pasted_obj)
  }
  return("not a character")
}

```


EXTENSION- PUZZLES

3.1

Write code that prints a multiplication table for numbers up to 12. i.e. there will be 12 rows and 12 columns, and each element should be the row number multiplied by the column number.

```{r}
matrix(1:12) %*% t(matrix(1:12))

```

3.2

Write a function that takes a vector, and creates a text art histogram, like the one below.

You may want to use the function cat(). This is similar to print(), but prints directly, without vector markers. You can also include newlines "\n" when using cat().

```{r}
make_histogram <- function(x){
  
  distinct <- sort(unique(x))
  
  for (i in distinct){
    
    count <- sum(x == i)
    
    cat(i)
    cat(" | ")
    cat(rep("*", count))
    cat("\n")
    
  }
  
}
```


3.3


Write a guessing game where the user has to guess a secret number. After every guess the program tells the user whether their number was too large or too small. At the end the number of tries needed should be printed.

To take user input, use the function readline(). Whenever R encounters readline, the code will stop, and the value assigned from readline will depend on the value that the user inputs. For more information see the help file.

```{r}
number <- round(runif(1, 1, 10))

guessed <- FALSE
attempts <- 0

while (!guessed){
  
  print("Guess a number between 1 and 10!")
  
  guess <- readline("")
  
  if (guess > number){
    
    print("Guess is too high")
    
    attempts <- attemps + 1
    
  }
  
  if (guess < number){
    
    print("Guess too low")
    
    attempts <- attempts + 1
    
  }
  
  if (guess == number){
    
    print("Guess is right! Well done!")
    
    print(paste("Attempts:", attempts))
    
    guessed <- TRUE
    
  }
  
}

```

3.4

By considering the terms in the Fibonacci sequence whose values do not exceed four million, find the sum of the even-valued terms.

You may want to read up on the definition of the Fibonacci sequence

```{r}
n_0 <- 0
n_1 <- 1

sum <- 0

while(n_1 <= 4000000){
  
  n_1 <- n_0 + n_1
  n_0 <- n_1 - n_0 # old n_1 value
  
  if (n_1 %% 2 == 0){
    sum <- sum + n_1
  }
  
}

sum
```

