---
title: "R Notebook"
output: 
  html_notebook: 
    fig_height: 5
    fig_width: 8
editor_options: 
  chunk_output_type: inline
---

In all plots you make, ensure your labels are appropriate. You also must write down a short summary of what each plot tells you in order to meet MVP.

Q1. 

```{r}

library(CodeClanData)
library(ggplot2)

```

```{r}
backpack

```


```{r}
# This gives the dimesnsions and names of this dataset
dim(backpack)
names(backpack)

```

Q2. 
What is the distribution of backpack weights that students are carrying? Choose an appropriate graph to show the distribution.

```{r}
ggplot(backpack) +
  geom_bar(aes(x = backpack_weight, fill = body_weight), position = "dodge") +
  scale_fill_brewer()

```

This graph shows that most students carried between 5 -10 (LBS) in backpack_weight

Q3. Are there differences in backpack weight between sex?

```{r}
ggplot(backpack) +
  geom_bar(aes(x = backpack_weight, fill = sex)) +
  scale_fill_brewer(palette = 7) +
  labs(
    x = " Backpack Weight",
    y = " Count",
    fill= "Sex"
  )
  

```

This graph shows that more females carried 5-10 (lbs) of backpack weight than males, 


q4. What is the relationship between body weight and backpack weight?


```{r}
ggplot(backpack) +
  geom_jitter(aes(x = backpack_weight, y = body_weight), colour = "blue", height = 5, width = 10)

```




Q.5 Is this relationship different between men and women? Choose an appropriate plot and interpret your findings.

```{r}
ggplot(backpack) +
  geom_line(aes(x = body_weight, y = backpack_weight, colour = sex)) 
  

```

The relationship between body weight and back_pack weight shows that in females in the smaller body weight -110-170 carried more bags beiween 10-20 or over. The males were between 150-200 in body weight

Q6. What proportion of students report back pain?

```{r}
ggplot(backpack) +
  geom_col(aes(x = year, y = back_problems), fill = "Light Green")

```

This graph shows that the proportion of year 3 students reported back pain, which means that younger and older students did not have as many problems

Q7 Is this proportion different between men and women?

```{r}
ggplot(backpack) +
  geom_col(aes(x = year, y = back_problems, fill = sex))

```
  
Again this graph shows that year 3 female students had more back problems than males. its also shows that for year 2 and year 6 only females had back problems, so the largest proportion of students with back problems were female

Q8. Now you’ve investigated your data separately, create plots to describe the relationship between backpack weight, body weight, gender and reported backpain. Try both backpack_weight and body_weight as well as the ratio.



```{r}

ggplot(backpack) +
  geom_histogram(aes(x = body_weight, fill = sex), col = "white") +
  facet_grid(back_problems ~ sex)

```
 
This graph shows the females and males that did and didn't have back problems, and their body weight

```{r}

ggplot(backpack) +
  geom_histogram(aes(x = backpack_weight, fill = sex), col = "white") +
  facet_grid(back_problems ~ sex)


```


This graph shows the females and males that did and didn't have back problems, and their backpack weight

