---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(lubridate)
library(tsibble)
library(tsibbledata)
library(fable)


```

1 Question 1
Load in the nyc_bikes data from the tsibbledata package. Have an initial look at it to see what you’re working with. Create three new columns: one that stores only the year the bike was used, one that stores only the month the bike was used, and one that stores the date. Use the data stored in start_time to create these new columns.

```{r}

nyc_bikes
```

```{r}
nyc_bikes_time <- nyc_bikes %>%
  mutate(year = year(start_time),
         month = month(start_time),
         date = as_date(start_time))

nyc_bikes_time
```

2 Question 2
Summarise the number of bike hire counts by month. Make a plot of this data. *Hint: remember that to group time series (tsibble) data, you need to use index_by instead of group_by before your summarise function. What does this plot tell you about the time series? Do you think this downsampled data would be adequate to build a forecast with?

```{r}
nyc_count <- nyc_bikes_time %>%
  group_by_key(bike_id) %>%
  index_by(month) %>%
  summarise(bike_hire_counts = n())
```


```{r}
ggplot(data = nyc_count, aes(x = month, y = bike_hire_counts, group = bike_id)) +
  geom_line(aes(color=bike_id))

```


3 Question 3
Now Summarise the number of bike hire counts by date. Make a plot of this new aggregated data. What does this plot tell you about the time series? Would this data be preferrable for time series forecasting compared to the monthly data?

```{r}
nyc_bikes_date <- nyc_bikes_time %>% 
  index_by(date) %>%
  summarise(bike_hire_counts = n())

nyc_bikes_date 
```


```{r}
ggplot(data = nyc_bikes_date, aes(x = date, y = bike_hire_counts)) +
  geom_line()


```

4 Question 4

Let’s begin to build a model. For this, we will use the downsampled by date dataset we created above in question 3. If you haven’t managed to get the code to do this, you can find it below.

```{r}

count <- nyc_bikes_date %>% 
  select(date, bike_hire_counts)

count %>%
  autoplot()
```

```{r}
fit_bikes <- nyc_bikes_date %>%
    fill_gaps() %>%
    model(
    snaive = SNAIVE(count),
    mean_model = MEAN(count),
    arima = ARIMA(count)
  )

```



```{r}
# fill the gaps : here i'll use median as it's a decent summary of how bike rentals work (not overly affected by seasonal changes)
nyc_bikes_filled <- nyc_bikes_date %>%
  fill_gaps(bike_hire_counts = as.integer(median(bike_hire_counts)))

```


```{r}

fit_bikes <- nyc_bikes_filled %>%
    model(
    snaive = SNAIVE(count),
    mean_model = MEAN(count),
    arima = ARIMA(count)
  )

```

