---
title: "R Notebook"
output: html_notebook
---

Create a vector called stock_prices with the following data points: 23, 27, 23, 21, 34, 24, 25, 30, 37, 30.

Create a vector called day_of_week, which repeats Monday to Friday twice. i.e. ‘Monday’, ‘Tuesday’, ‘Wednesday’, ‘Thursday’, ‘Friday’, ‘Monday’, ‘Tuesday’, ‘Wednesday’, ‘Thursday’, ‘Friday’.

```{r}
stock_prices <- c(23, 27, 23, 21, 34, 24, 25, 30, 37, 30)
```


```{r}
days_of_the_week <-  rep(c("monday", "tuesday", "wednesday", "thursday", "friday"), each = 2)

```
```{r}
days_of_the_week

```

```{r}
stock_prices_2 <- c(23, 27, 23, 21, 34, 24, 25, 30, 37, 30)
day_of_week <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday', 'Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday')
```


Create a vector called day_of_week_factor, which has the same data as day_of_week, but is of factor type.

Created a vector called named_stock_prices, which contains the stock prices, named by days of the week.

```{r}
days_of_the_week_factor <- factor(c(days_of_the_week))
```
```{r}
days_of_the_week_factor
```


```{r}
days_of_the_week
```


```{r}
named_stock_prices <-c(stock_prices, days_of_the_week)
```

Solution

```{r}
day_of_week_factor <- as.factor(day_of_week)
```

```{r}

named_stock_prices_2 <- stock_prices_2
names(named_stock_prices_2) <- day_of_week
```

Make a data frame called stock_data containing the stock price and day of the week information, and View() the result.

```{r}
data.frame(stock_prices, days_of_the_week)
```

```{r}
stock_data <- data.frame(stock_prices, days_of_the_week)
```

```{r}
stock_data
```


```{r}
view(stock_data)
```


```{r}
View(stock_data)
```

Solution

```{r}
stock_data <- data.frame(stock_prices = stock_prices, day_of_week = day_of_week)
```

```{r}
View(stock_data)

```

Make a list called stock_list with named elements as follows: the vector of stock_prices, the vector day_of_week, and the data frame you made in question 3. Choose sensible names for the list elements.

```{r}
stock_list <- list(stock_prices, days_of_the_week, stock_data)

```

```{r}
stock_prices[[4]][[5]][[6]]
```

```{r}
stock_prices$4$5$
```

SOLUTION

```{r}
stock_list <- list(stock_prices = stock_prices, day_of_week = day_of_week, stock_data = stock_data)
```

In this part, you will practice selecting elements from vectors, data frames, matrices and lists.

Select the 4th, 5th and 6th stock prices from stock_prices.

Select the stock prices greater than 30 from stock_prices.

```{r}
stock_prices_2 [4:6]
```

```{r}
stock_prices_2 [stock_prices_2 > 30]
```

Select the day_of_week column from the stock_data data frame.

```{r}
stock_data$day_of_week

```


Select the rows of the stock_data data frame where the stock_prices are less than 25.

```{r}

subset(stock_data, stock_prices < 25)

```

Select the 1st column of stock_data.

```{r}
stock_data[ ,1]
```


Select the 3rd row of stock_data.

```{r}
stock_data [3,]
```


Select the 3rd element in the 1st column of stock_data.

```{r}
stock_data [3,1]
```

Select the data frame from within the list of stocks.

```{r}
stock_list$stock_data

```

Extension

```{r}
install.packages("tidyquant")
```

```{r}
library(tidyquant)
```

Use it to get stock prices from a company of your choice from the current year. You’ll need to look up the help file to do this.

```{r}
google_prices  <- tq_get("GOOGL", get = "stock.prices", from = "2020-01-01")

```

edit the data frame to keep only the date and close columns.

```{r}
google_prices[ , c("date", "close")]

```

Subset to find the rows at which the close price is higher than average.

```{r}

subset(google_prices, close > mean(close))
```

